---
title: "Interactive Data Visualization with Plotly"
author: "Misha Ash"
date: "05/25/2018"
output: html_document
---

```{r setup, include = FALSE, echo = TRUE, cache = TRUE, fig.align = 'center'}

knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(ggplot2)
library(plotly)

library(ggthemes)
#library(magrittr)
library(tidyverse)
```

```{r setup_output, include=FALSE, eval = FALSE}
# unsuccessfully tried to use this as a means of adjusting vertical space between chunks and text

hook_output_def = knitr::knit_hooks$get('output')
knitr::knit_hooks$set(output = function(x, options) {
  if (!is.null(options$vspaceout)) {
    end <- paste0("\\vspace{", options$vspaceout, "}")
    stringr::str_c(hook_output_def(x, options), end)
  } else {
    hook_output_def(x, options)
  }
})
```

```{r load data, echo = FALSE, include = FALSE}

# ELAN encoding (n=24) with added accuracy, item code, and other item variables
study_01 <- readr::read_csv("~/R/CGChains/data/CGChains_CODED-ELAN-DATA_2018-04-24.csv")

```

```{r recode elan data, include = FALSE}
# encode gesture presence and spatiality as dichotomous indicator variables

study_01 <- study_01 %>%
  
  # encode as data frame
  #as_data_frame() %>%
  
  # add dichotomous variables
  mutate(GESTURE_PRESENT = if_else(study_01$`gesture_presence` == "yes", 1, 0)) %>%
  mutate(SPATIAL_TRIAL = if_else(study_01$trial_type == "spatial", 1, 0)) %>%
  mutate(VERTICAL_TRIAL = if_else(study_01$item_axis == "vertical", 1, 0)) %>%
  
  # add difficulty variable based on deSoto (1965)
  mutate(deSoto_difficulty = cut(x = correct_deSoto_without_invalid, breaks = c(38.3, 41.5, 60.5, 61.8, 100),
                                 right = FALSE, labels = c("HARDEST", "HARD", "EASY", "EASIEST")))
```


```{r variable types lists, include = FALSE}
# variable types list
study_01_VARs <- study_01 %>% 
  summarize_all(typeof) %>% 
  gather()

```

```{r gesture_type_axis subset, include = FALSE}
# select variables of interest
gesture_type_axis <- study_01 %>%
  dplyr::select(SUBJ, LIST, trial_number, ITEM_CODE, SPATIAL_TRIAL, VERTICAL_TRIAL, SPATIAL_FIRST, 
         GESTURE_PRESENT, gesture_axis, ACCURACY, 
         difficulty_indicator_without_invalid, correct_deSoto_without_invalid, deSoto_difficulty)

```

```{r gesture_type_axis_coded factor encodings, include = FALSE, eval = FALSE}
# encoding trial spatiality and axis as factors 

gesture_type_axis_coded <- gesture_type_axis
gesture_type_axis_coded$SPATIAL_TRIAL <- factor(gesture_type_axis_coded$SPATIAL_TRIAL, 
                                                levels = c(0, 1), 
                                                labels = c("nonspatial", "spatial"))

gesture_type_axis_coded$VERTICAL_TRIAL <- factor(gesture_type_axis_coded$VERTICAL_TRIAL, 
                                                levels = c(0, 1), 
                                                labels = c("horizontal", "vertical"))

#gesture_type_axis_coded$GESTURE_PRESENT <- factor(gesture_type_axis_coded$GESTURE_PRESENT, 
 #                                               levels = c(0, 1))

# variable types list
gesture_type_axis_coded_VARs <- gesture_type_axis_coded %>% 
  summarize_all(typeof) %>% 
  gather()

```

```{r mean proportion of gesture production by trial type, include = FALSE}
prop_gesture_type_axis <- study_01 %>%
  dplyr::select(GESTURE_PRESENT, trial_type, item_axis) %>%
  dplyr::group_by(trial_type, item_axis) %>%
  dplyr::summarize(prop_gesture = sum(GESTURE_PRESENT)/240) 
```

```{r subj means by trial type and axis, include = FALSE}
# participant means by type and axis conditions
prop_subj_gesture_type_axis <- study_01 %>%
  dplyr::select(SUBJ, GESTURE_PRESENT, trial_type, item_axis) %>%
  dplyr::group_by(SUBJ, trial_type, item_axis) %>%
  dplyr::summarize(prop_gesture = sum(GESTURE_PRESENT)/10) 
```

```{r level 1 data, include = FALSE}
# no interaction terms
level_01_data <- study_01 %>%

# INDEX VARs
  
  # ITEM: individual stimulus problem numbers
  mutate(ITEM = as.integer(ITEM_CODE)) %>%
  # LIST
  mutate(LIST = LIST) %>%
  # TRIAL
  mutate(TRIAL = trial_number) %>%
  # SUBJECT ID
  mutate(ID = SUBJ) %>%

# PRIMARY IVs

  # SPATIALITY: spatial trials = 1, non-spatial trials = 0
  mutate(SPACE_IV = if_else(study_01$trial_type == "spatial", 1, 0)) %>%
  # AXIS: vertical trials = 1, horizontal trials = 0
  mutate(AXIS_IV = if_else(study_01$item_axis == "vertical", 1, 0)) %>%
  # DIFFICULTY: difficult trials = 2, easy trials = 1, invalid trials = 0
  mutate(DIFF_IV = if_else(study_01$DIFFICULTY_with_invalid == "HARD", 2,
                      if_else(study_01$DIFFICULTY_with_invalid == "EASY", 1, 0))) %>%

  # ORDER: spatial first = 1, non-spatial first = 0
  mutate(ORDER = as.integer(if_else(study_01$SPATIAL_FIRST == 1, 1, 0))) %>%
  
# SECONDARY IVs ## leaving this out for now
  # difficulty based on deSoto (1965): HARDEST = 4, HARD = 3, EASY = 2, EASIEST = 1
  #mutate(deSoto_difficulty_IV2 = cut(x = correct_deSoto_without_invalid, breaks = c(38.3, 41.5, 60.5, 61.8, 100),
                                 #right = FALSE, labels = c(4, 3, 2, 1))) %>%
  
# DVs
  # GESTURE_PRESENCE: present = 1, absent = 0
  mutate(GESTURE_DV = if_else(study_01$gesture_presence == "yes", 1, 0)) %>%
  # GESTURE_AXIS: lateral = 1, vertical = 2, saggital = 3, other/unclear = 4, ERROR = 5
  mutate(AXIS_DV = if_else(study_01$gesture_axis == "lateral", 1, 
                   if_else(study_01$gesture_axis == "vertical", 2,
                   if_else(study_01$gesture_axis == "sagittal", 3, 
                   if_else(study_01$gesture_axis == "other/unclear", 4, 5))))) %>%

  # ACCURACY: correct = 1, incorrect = 0
  mutate(ACCURACY_DV = if_else(study_01$ACCURACY == 1, 1, 0)) %>%
  
# SELECT VARIABLES

  dplyr::select(ITEM, LIST, TRIAL, ID, 
         SPACE_IV, AXIS_IV, DIFF_IV, 
         ORDER, # deSoto_difficulty_IV2, 
         GESTURE_DV, AXIS_DV, ACCURACY_DV)
```

```{r X gesture & accuracy by subject, include = FALSE}
# calculate gesture and accuracy proportions by subject
X <- level_01_data %>%
  #dplyr::select(SUBJ, GESTURE_DV, ACCURACY_DV) %>%

  dplyr::group_by(ID, SPACE_IV) %>%
  dplyr::summarize(gestures = sum(GESTURE_DV), 
                   percent_gestured = gestures / (length(unique(ITEM))*2),
                   gesture_se = sd(GESTURE_DV)/sqrt((length(unique(ITEM)))*2),
                   
                   errors = length(unique(ITEM)) - sum(ACCURACY_DV),
                   percent_errors = errors / (length(unique(ITEM))*2),
                   error_se = sd(1 - ACCURACY_DV)/sqrt((length(unique(ITEM)))*2)
                   ) %>%

  as_factor(ID)
```


I used this assignment as an opportunity to experiment with some `Plotly` features to visualize data in a clean way while maintaining access to precise values with `hoverlabel` and `tooltip`.

#### bar plot distributions

The following plotly histogram shows the proportion of gestures produced by each participant. Hovering over a bar shows its precise y value.

```{r echo = FALSE, message = FALSE}
ttl_gest_prop <- prop_subj_gesture_type_axis %>%
  mutate(prop = prop_gesture/4) %>%
  dplyr::select(SUBJ, prop) %>%
  group_by(SUBJ) %>%
  summarize(prop_gesture = sum(prop)) #%>%
  #tidyr::spread(key = trial_type, value = prop_gesture)

plotly_gest_histo_01 <- plot_ly(ttl_gest_prop,
                                x = ~SUBJ, y = ~prop_gesture,
                                hoverinfo = 'y',
                                type = 'bar', width = 0.5, 
                                marker = list(color = "#a2adba"),
                                opacity = .7) %>%
  layout(title = "gesture production",
         yaxis = list(title = 'proportion of trials with gesture'), 
         barmode = 'stack',
         xaxis = list(
         title = "participant number", range = c(13, 38),
         autotick = FALSE, 
         ticks = "outside",
         tick0 = 14, # can't seem to make it start from the second value; changed range instead
         dtick = 1,
         ticklen = 3,
         tickwidth = 1,
         tickcolor = "black"),
         margin = list(b = -10, pad = 1))
           

plotly_gest_histo_01
```

* design choices
  - enhanced aesthetic by changing default fill color and opacity of the bars
  - adjusted tick marks on x axis to clearly show correspondence of values with bars
  - moved x-axis label down (was too close to axis)
  - set the range of x-axis so leftmost and rightmost bars do not run into sides
  

* issues
  - The aforementioned range manipulation was done because by default the minimum and maximum x values caused the corresponding bars to run into the sides. I could not remove the first and last labels of the x-axis, however, having tried the `tick0` parameter. 

This distribution can be further enhanced in a few ways. One is use the color channel to distinguish gestures along another variable of interest---spatiality in this case:

```{r gesture production histogram, echo = FALSE}
# munge a tidy frame for stacked histogram
gest_spatiality <- prop_subj_gesture_type_axis %>%
  mutate(prop = prop_gesture/4) %>%
  dplyr::select(SUBJ, trial_type, prop) %>%
  group_by(trial_type, SUBJ) %>%
  summarize(prop_gesture = sum(prop)) %>%
  spread(key = trial_type, value = prop_gesture)

# rename column because plotly appears not to like col names with dash and quoted names interpreted as strings
names(gest_spatiality)[names(gest_spatiality) == 'non-spatial'] <- 'abstract' 


plotly_gest_histo_02 <- plot_ly(gest_spatiality, 
                                  x = ~SUBJ, y = ~spatial, type = 'bar', name = 'spatial', 
                                width = 0.5, marker = list(color = "#66c2a5"), opacity = 0.7) %>%
  add_trace(gest_spatiality, y = ~abstract, name = 'non-spatial',
            marker = list(color = "#8da0cb"), opacity = 0.5) %>% #"#a2adba")) %>%
  
  layout(title = "gesture production",
         yaxis = list(title = 'proportion of trials with gesture'), 
         barmode = 'stack',
         xaxis = list(
         title = "participant number", range = c(13, 38),
         autotick = FALSE, 
         ticks = "outside",
         tick0 = 14, # can't seem to make it start from the second value; changed range instead
         dtick = 1,
         ticklen = 3,
         tickwidth = 1,
         tickcolor = "black"),
         margin = list(b = -10, pad = 1))

plotly_gest_histo_02

```

Although hovering shows the relative proportion of trials with gesture split according to the non/spatial variable, the it is difficult to discern whether there is an overall pattern. One reason might be the number of bars that are split nearly equally, making the difference nearly indiscernable without looking at the exact values. Another reason is that the current arrangement of information is not amenable to a general comparison across all x values.

To attempt to account for the latter limitation, the bars are rearranged in order of ascending value (while keeping the labels constant so correspondence between participant number and x-axis label is maintained). I did this while also trying out `ggplotly` to translate a `ggplot2` object into a `plotly` object:

```{r ggplotly hist 1, echo = FALSE, fig.width=6.5, fig.height=3.5}

dist_gest <- X %>%
  
  ggplot() +

  geom_col(aes(x = reorder(ID, percent_gestured), y = percent_gestured, #group = SUBJ, 
                 fill = as.factor(SPACE_IV), alpha = .85)) +
  
   # scale_y_continuous(expand = c(0, 0)) +
   # scale_x_continuous(breaks = c(seq(1:24)), expand = c(0, 0)) +
  
    # change color scheme of fill colors
    scale_fill_manual("fill_constant", values = c("#8da0cb", "#66c2a5")) +
  
  scale_y_continuous(expand = c(0, 0)) +

    theme_light() +
    theme(panel.grid.major.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.y = element_blank(), # remove minor horizontal grid lines
        # remove legends (unnecessary here)
        legend.position = "none", ### TEMP
        plot.title = element_text(margin = margin(t = 0, r = 0, b = 10, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = -5, l = 0)), 
        panel.border = element_blank(), # remove outer panel border
        axis.ticks.y = element_blank(), # remove ticks from y axis
        text = element_text(family = "Arial")) +
  
    ggtitle('gesture production by participant') +
    xlab('participant id (arranged by total gesture production)\n') +
    ylab('proportion of trials with gesture')

ggplotly(dist_gest, tooltip = c("x", "y", "SPACE_IV"), session = "knitr")

```

`r "\n"`
This provides a better sense of the distribution, including the relative proportions of spatial and non-spatial trials (we can already see that spatial trials appear to account for more of the gestures overall). Lastly, another potentially useful addition is a *horizonal line* indicating the mean of the distribution:


```{r ggplotly hist 3, echo = FALSE, fig.width=6.5, fig.height=3.5, warning = FALSE}

'total mean =' <- round(mean(X$percent_gestured)*2, digits = 3)

dist_gest_03 <- X %>%
  
  ggplot() +
    
   geom_hline(aes(yintercept = mean(percent_gestured)*2, 'total mean' = 0.524), 
              color = "#2e184b", size = 0.3) +
  
    # label mean line
    geom_text(x = 3, y = (mean(X$percent_gestured)*2 + 0.03), 
              aes(label = "total mean = "), color = "#2e184b") +
    
    geom_text(x = 4.75, y = (mean(X$percent_gestured)*2 + 0.03), 
              aes(label = round(mean(percent_gestured)*2, digits = 3)), color = "#2e184b") +
  
    geom_col(aes(x = reorder(ID, percent_gestured), y = percent_gestured, #group = SUBJ, 
                 fill = as.factor(SPACE_IV), alpha = .85)) +
  
    # change color scheme of fill colors
    scale_fill_manual("fill_constant", values = c("#8da0cb", "#66c2a5")) +
  
  scale_y_continuous(expand = c(0, 0)) +
   # scale_x_continuous(breaks = c(seq(1:24)), expand = c(0, 0)) +
  
    theme_light() +
    theme(panel.grid.major.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.y = element_blank(), # remove minor horizontal grid lines
        # remove legend
        legend.position = "none",
        plot.title = element_text(margin = margin(t = 0, r = 0, b = 10, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)), 
        panel.border = element_blank(), # remove outer panel border
        axis.ticks.y = element_blank(), # remove ticks from y axis
        text = element_text(family = "Arial")) +
  
    ggtitle('gesture production by participant') +
    xlab('participant id (arranged by total gesture production)\n') +
    ylab('proportion of trials with gesture')

ggplotly(dist_gest_03, tooltip = c("total mean", "spatial mean", "percent_gestured", "SPACE_IV"))

```

This provides a nice reference point, though the labeling does not render well and could use more refining. Or, since these are `plotly` graphics, the lines can be placed and labels can be based on hovering (group mean lines also added):

```{r ggplotly hist 4, echo = FALSE, fig.width=6.5, fig.height=3.5, warning = FALSE}


dist_gest_04 <- X %>%
  
  ggplot() +
    
  # mean gesture production line
#  geom_hline(aes(yintercept = mean(percent_gestured)*2), color = "#2e184b", size = 0.3) +
  
      geom_hline(aes(yintercept = mean(percent_gestured)*2, 'total mean' = 0.524), 
              color = "#2e184b", size = 0.3) +

 # label mean line
#    geom_text(x = 3, y = (mean(X$percent_gestured)*2 + 0.03), 
#              aes(label = "total mean = "), color = "#2e184b") +
#    
#    geom_text(x = 4.75, y = (mean(X$percent_gestured)*2 + 0.03), 
#              aes(label = round(mean(percent_gestured)*2, digits = 3)), color = "#2e184b") +
  
    geom_col(aes(x = reorder(ID, percent_gestured), y = percent_gestured, #group = SUBJ, 
                 fill = as.factor(SPACE_IV), alpha = .85)) +
  
  
### group means... cluncky code: there must be a simpler way!
  
  # spatial mean line
  geom_hline(aes(yintercept = 
      (sum(X[X$SPACE_IV == 1, "percent_gestured"]) / 
      tally(X[X$SPACE_IV == 1, "percent_gestured"]))[1,1], 'spatial mean' = 0.305),
      color = "#5bae94") +
  
  # spatial mean labels  
#  geom_text(x = 3, 
#            y = (sum(X[X$SPACE_IV == 1, "percent_gestured"]) / 
#              tally(X[X$SPACE_IV == 1, "percent_gestured"])[1,1] + 0.03)[1,1], 
#            aes(label = "spatial mean = "), color = "#5bae94") +
  
#  geom_text(x = 7, 
#            y = (sum(X[X$SPACE_IV == 1, "percent_gestured"]) / 
#              tally(X[X$SPACE_IV == 1, "percent_gestured"])[1,1] + 0.03)[1,1], 
#            
#            aes(label = round((sum(X[X$SPACE_IV == 1, "percent_gestured"]) / 
#      tally(X[X$SPACE_IV == 1, "percent_gestured"]))[1,1], digits = 3)), color = "#5bae94") +
  
  # non-spatial mean labels  
#  geom_text(x = 3.5, 
#            y = (sum(X[X$SPACE_IV == 0, "percent_gestured"]) / 
#              tally(X[X$SPACE_IV == 0, "percent_gestured"])[1,1] + 0.03)[1,1], 
#            aes(label = "non-spatial mean = "), color = "#5bae94") +
  
#  geom_text(x = 7.5, 
#            y = (sum(X[X$SPACE_IV == 0, "percent_gestured"]) / 
#              tally(X[X$SPACE_IV == 0, "percent_gestured"])[1,1] + 0.03)[1,1], 
            
#            aes(label = round((sum(X[X$SPACE_IV == 0, "percent_gestured"]) / 
#      tally(X[X$SPACE_IV == 0, "percent_gestured"]))[1,1], digits = 3)), color = "#5bae94") +
  
  
#  geom_text(x = 6, y = 0.554, aes(label = round(mean(percent_gestured)*2, digits = 3)), 
#                                  color = "#2e184b") +
  
    # non-spatial mean
  geom_hline(aes(yintercept = 
      (sum(X[X$SPACE_IV == 0, "percent_gestured"]) / 
      tally(X[X$SPACE_IV == 0, "percent_gestured"]))[1,1]), 'non-spatial mean' = 0.219,
      color = "#7e90b6") +
  
#  geom_hline(aes(yintercept = mean((X[X$SPACE_IV == 0, "percent_gestured"])[,])), color = "#FFA319") +
#  geom_hline(aes(yintercept = mean((X[X$SPACE_IV == 1, "percent_gestured"])[,])), color = "#FFA319") +
  

### temporarily remove ribbon (need to fix error margin)
 #   annotate('ribbon', x = c(-Inf, Inf), 
  #           ymin = summary_subj_err_gest$mean_percent_gestured - summary_subj_err_gest$mean_gesture_se, 
   #          ymax = summary_subj_err_gest$mean_percent_gestured + summary_subj_err_gest$mean_gesture_se, 
    #         alpha = 0.2, fill = "#628B9F") +
  
  geom_col(aes(x = reorder(ID, percent_gestured), y = percent_gestured, #group = SUBJ, 
                 fill = as.factor(SPACE_IV), alpha = .85)) +
  

  
   # scale_y_continuous(expand = c(0, 0)) +
   # scale_x_continuous(breaks = c(seq(1:24)), expand = c(0, 0)) +
  
    # change color scheme of fill colors
    scale_fill_manual("fill_constant", values = c("#8da0cb", "#66c2a5")) +
  
  scale_y_continuous(expand = c(0, 0)) +
   # scale_x_continuous(breaks = c(seq(1:24)), expand = c(0, 0)) +
  
    theme_light() +
    theme(panel.grid.major.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.x = element_blank(), # remove vertical grid lines
        panel.grid.minor.y = element_blank(), # remove minor horizontal grid lines
        # remove legends (unnecessary here)
        legend.position = "none", ### TEMP
        plot.title = element_text(margin = margin(t = 0, r = 0, b = 10, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)), 
        axis.title.x = element_text(margin = margin(t = 10, r = 0, b = 0, l = 0)), 
        panel.border = element_blank(), # remove outer panel border
        axis.ticks.y = element_blank(), # remove ticks from y axis
        text = element_text(family = "Arial")) +
  
    ggtitle('gesture production by participant') +
    xlab('participant id (arranged by total gesture production)\n') +
    ylab('proportion of trials with gesture')

ggplotly(dist_gest_04, tooltip = c("total mean", "spatial mean", "spatial mean", "non-spatial mean", 
                                   "percent_gestured", "SPACE_IV"))

```

Having the extra mean lines, though potentially informative, seems to be excessive as it crowds the graphic. This brings me to the task of visualizing means while attempting to maintain keep the distribution in view.

#### the split violin

A potential alternative to the bar chart, which is traditionally used to visualize group means, we can instead simultaneously show the mean and the distribution of the group from which the mean is calculated by using a violin plot. The standard violin plot is symmetrical, but to visualize two groups it may be more beneficial to split the violin, as follows:

```{r violin 1, warning = FALSE, echo = FALSE}

gesture_proportion_violin <- prop_subj_gesture_type_axis %>%
  
  plot_ly(type = 'violin') %>%
  add_trace(
    x = ~ item_axis[prop_subj_gesture_type_axis$trial_type == 'spatial'],
    y = ~ prop_gesture[prop_subj_gesture_type_axis$trial_type == 'spatial'],
    
    legendgroup = 'spatial',
    scalegroup = 'spatial',
    name = 'spatial',
    side = 'negative', # left side
    opacity = 0.8,
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    line = list(
      color = '#66c2a5' # selected from qualitative colorblind safe scheme from Color Brewer
    )
  ) %>%
  
  add_trace(
    x = ~ item_axis[prop_subj_gesture_type_axis$trial_type == 'non-spatial'],
    y = ~ prop_gesture[prop_subj_gesture_type_axis$trial_type == 'non-spatial'],
    
    legendgroup = 'non-spatial',
    scalegroup = 'non-spatial',
    name = 'non-spatial',
    side = 'positive', # right side
    opacity = 0.7,
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    line = list(
      color = '#8da0cb' # selected from qualitative colorblind safe scheme from Color Brewer
    )
  ) %>% 
  layout(
    title = "Gesture Production by Trial Type<br><i>(proportion observations are grouped by subject)",
    xaxis = list(
      title = ""  
    ),
    yaxis = list(
      title = "proportion of trials with gesture",
      zeroline = F
    ),
    violingap = 0,
    violingroupgap = 0.1,
    violinmode = 'overlay', # because violins are split they need to be overlaid
    tracegroupgap = 1
  )

gesture_proportion_violin
```

Although it is a nice first attempt, this visualization has one particularly glaring problem: The distribution exceeds the range of values in the dataset! This is due to the smooting kernel density function and can be addressed by limiting the range of values, which effectively sets cutoff points for the plot. We also remove the inner box plots, which add clutter and instead place the crucial part of the information they add, the mean, as a dashed line within each distribution:

```{r violin 2, warning = FALSE, echo = FALSE}

gesture_proportion_violin <- prop_subj_gesture_type_axis %>%
  
  plot_ly(type = 'violin') %>%
  add_trace(
    x = ~ item_axis[prop_subj_gesture_type_axis$trial_type == 'spatial'],
    y = ~ prop_gesture[prop_subj_gesture_type_axis$trial_type == 'spatial'],
    
    legendgroup = 'spatial',
    scalegroup = 'spatial',
    name = 'spatial',
    side = 'negative', # left side
    opacity = 0.8,
    box = list(
      visible = F
    ),
    meanline = list(
      visible = T
    ),
    line = list(
      color = '#66c2a5' # selected from qualitative colorblind safe scheme from Color Brewer
    )
  ) %>%
  
  add_trace(
    x = ~ item_axis[prop_subj_gesture_type_axis$trial_type == 'non-spatial'],
    y = ~ prop_gesture[prop_subj_gesture_type_axis$trial_type == 'non-spatial'],
    
    legendgroup = 'non-spatial',
    scalegroup = 'non-spatial',
    name = 'non-spatial',
    side = 'positive', # right side
    opacity = 0.7,
    box = list(
      visible = F
    ),
    meanline = list(
      visible = T
    ),
    line = list(
      color = '#8da0cb' # selected from qualitative colorblind safe scheme from Color Brewer
    )
  ) %>% 
  layout(
    title = "Gesture Production by Trial Type<br><i>(proportion observations are grouped by subject)",
    xaxis = list(
      title = ""
    ),
    yaxis = list(
      title = "proportion of trials with gesture",
      zeroline = F,
      range = c(0, 1)
    ),
    violingap = 0,
    violingroupgap = 0.1,
    violinmode = 'overlay', # because violins are split they need to be overlaid
    tracegroupgap = 1,
    margin = list(
  l = -1,
  r = 0,
  b = -1,
  t = -5,
  pad = 0
)
  )

gesture_proportion_violin
```

#### the story

These experiments explore some of the potential of alternative ways to display scientific data. While priliminary, they show that alternatives exist that can present data in more informative and engaging ways. Interactivity has some uses here, allowing the audience to explore the information by homing in on data, looking at specific data points and, in situations not shown here, using additional interactive features of `plotly` such as zoom. This is a double-edged sward, however, and care should be taken to visualize scientific results in alternative ways only when there is a good justification for doing so. Labelling and other interactive elements may be excessive or unnecessary, and in many cases traditional forms may be preferable precisely for the reasons that made them the standard in the first place.



```{r box plot 1, warning = FALSE, echo = FALSE}

prop_subj_gesture_type_axis$trial_type <- factor(
  prop_subj_gesture_type_axis$trial_type, levels = c("spatial", "non-spatial"))

p <- plot_ly(y = ~ prop_subj_gesture_type_axis$prop_gesture, 
             type = "box", boxpoints = "all", jitter = 0.3,
             pointpos = 0)

p <- plot_ly(prop_subj_gesture_type_axis, x = ~ item_axis,
             y = ~ prop_gesture, color = ~ trial_type, type = "box",
             boxpoints = "all", jitter = 0.3, pointpos = 0) %>%
  layout(boxmode = "group")

p

```

```{r excluded split violin, include = FALSE, eval = FALSE}

pointposMale <- c(-0.9,-1.1,-0.6,-0.3)
pointposFemale <- c(0.45,0.55,1,0.4)
showLegend <- c(T,F,F,F)

p <- plot_ly(type = 'violin')

i = 0
for (i in 1:2) {
  p <- add_trace(
    p,
    x = prop_subj_gesture_type_axis$item_axis[prop_subj_gesture_type_axis$trial_type == 'spatial' & 
                                           prop_subj_gesture_type_axis$item_axis == 
                                           unique(prop_subj_gesture_type_axis$item_axis)[i]],
    
    y = prop_subj_gesture_type_axis$prop_gesture[prop_subj_gesture_type_axis$trial_type == 'spatial' & 
                                           prop_subj_gesture_type_axis$item_axis == 
                                           unique(prop_subj_gesture_type_axis$item_axis)[i]],
    

    hoveron = "all",
    legendgroup = 'spatial',
    scalegroup = 'spatial',
    name = 'spatial',
    side = 'negative',

## haven't figured out how to differentially control opacity
#   group = prop_subj_gesture_type_axis$item_axis,
#    opacity = .5,

    box = list(
      visible = T
    ),
#    points = 'all',
#    pointpos = pointposMale[i],
#    jitter = 0,
#    scalemode = 'count',
    meanline = list(
      visible = F
    ),
    line = list(
      color = "#8dd3c7"
    ),
    marker = list(
      line = list(
        width = 1,
        color = "#8dd3c7"
      ),
      symbol = 'line-ns'
    ),
    showlegend = showLegend[i]
    ) %>% 
    
    
    add_trace(
    x = prop_subj_gesture_type_axis$item_axis[prop_subj_gesture_type_axis$trial_type == 'non-spatial' & 
                                           prop_subj_gesture_type_axis$item_axis == 
                                           unique(prop_subj_gesture_type_axis$item_axis)[i]],
    
    y = prop_subj_gesture_type_axis$prop_gesture[prop_subj_gesture_type_axis$trial_type == 'non-spatial' & 
                                           prop_subj_gesture_type_axis$item_axis == 
                                           unique(prop_subj_gesture_type_axis$item_axis)[i]],
      
      hoveron = "all",
      legendgroup = 'non-spatial',
      scalegroup = 'non-spatial',
      name = 'non-spatial',
      side = 'positive',
      box = list(
        visible = T
      ),
#      points = 'all',
#      pointpos = pointposFemale[i],
#      jitter = 0,
      scalemode = 'count',
      meanline = list(
        visible = F
      ),
      line = list(
        color = "#bebada"
      ),
      marker = list(
        line = list(
          width = 2,
          color = "#bebada"
        ),
        symbol = 'line-ns'
      ),
      showlegend = showLegend[i]
  )
}

p <- layout(
  p,
  title = "Proportion of problems on which gestures were produced<br><i>...",
  yaxis = list(
    zeroline = F
  ),
  violingap = 0,
  violingroupgap = 0.1,
  violinmode = 'overlay', # because violins are split they need to be overlaid
  legend = list(
    tracegroupgap = 1
  )
)

p

# htmlwidgets::saveWidget(as_widget(p), "violin_example.html")
```


